\section*{Problem 4: The Tully-Fisher Relation}

the Tully-Fisher relation, which states that the
luminosity L of a spiral galaxy is roughly proportional to the circular velocity $v_c$ of stars to the fourth power:

\begin{equation}
    L \propto v^4_c.
\end{equation}

This used to be used to infer distances on the cosmic distance ladder until better techniques arised, the Tully-Fisher relation has a $\sim 10\%$ scatter. 

\subsection*{Problem 4A}

The Virial Theorem states $2K + V = 0$, where $K$, and $V$ are the kinetic and potential energy respectfully. 

In our case we can write:
\begin{equation}
    Mv_c^2 + \frac{\alpha GM^2}{R} = 0
\end{equation}
here, M is the mass the galaxy, and R is the distance from the centre, $\alpha$ is some correction factor.

We can rewrite this as:
\begin{equation}
    v_c^2 \propto \frac{M}{R}
\end{equation}


For the case of a constant mass-luminosity ratio ($M/L = c_1$), we can replace $M$ in terms of $L$. 

\begin{equation} \label{problem_2_a}
    \boxed{v_c^2 \propto \frac{L}{R}}
\end{equation}

\subsection*{Problem 4B}
Here we now also assume constant surface brightness $c_2 = \frac{L}{A}$, where A is area. We can write $A = \alpha R^2$, where again $\alpha$ is some correction factor. Using our solution to problem 2A equation \ref{problem_2_a}, we can write:

\begin{equation}
    R \propto L^{1/2}
\end{equation}

and subsisting into equation \ref{problem_2_a},

\begin{equation}
    v_c^2 \propto L^{1/2}
\end{equation}

rewriting:
\begin{equation}
    \boxed{L \propto v_c^4}
\end{equation}


\section*{Problem 4C}
Assumptions made in this problem were 1) Virial Theorem holds, 2) constant Mass-to-Luminosity ratio, 3) constant surface brightness, 4) no darkmatter.


1) Virtualisation holds for bound systems like some atoms, globular cluster, (closed) planetary orbits. and for galaxies by the defintion given in class. 

2) is unrealistic unless there is no active star-formation. We know that for main sequence stars there $L \sim M^\alpha$, with $\alpha = 3.5$ - a power-law not a constant. 

3) unrealistic, galaxies are brighter at their center and dimmer at their edges

4) We never invoked the existence of Dark matter into our analysis, at least to my understanding the virtualisation I used is for normal matter.  

